---
title: "Statiscal Inference Final Project"
author: "Maria Cristina Alameda Salas"
date: "25/4/2023"
output: html_document
---




## Part 2: Basic Inferential Data Analysis


Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

### Loading the ToothGrowth data

```{r}

# load "tidyverse" plotting package
library(ggplot2)



# Load data
data(ToothGrowth)

```


### Performing some basic exploratory data analyses
In the following code, we can see that 3 columns compose our dataset: len, supp & dose.
```{r}
# Show first rows
head(ToothGrowth)
```


```{r}
dim(ToothGrowth)
```




### Provide a basic summary of the data.

This data contains 3 columns. The first column (*len*) talks about the length of tooth of guinea pigs. It is a continious value in range 4.20 - 33.90 with a mean of 18.81. The columns *supp* is about the supplement took by the animal. This column is discrete, with just two possible values (OJ & VC).

The last column (*dose*) shows data about dosage of supplement in mg. It goes from 0.5 mg to 2 mg.

```{r}
summary(ToothGrowth)
```

Let's plot the previous information by *supp* group.

```{r}
qplot(factor(supp), len, data = ToothGrowth, facets=~dose, 
      main="Tooth growth of guinea pigs by supplement type and dosage (mg)",
      xlab="Supplement type", ylab="Tooth length") + 
      geom_boxplot(aes(fill = supp)) +
      scale_fill_manual(values=c("#f9b4ab", "#679186"))
```